---
title: "Mining land use"
author: "Victor Maus"
date: "3/8/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(parallel)
library(stringr)
library(here)
library(sf)
library(viridis)
library(rnaturalearth)
knitr::opts_chunk$set(echo = TRUE)
```

## Load tables

```{r load-tables}
version <- "20210317"
cluster_type <- "hcluster"
derived_data_path <- here::here('analysis', 'data', 'derived_data', version)
derived_tables_path <- glue::glue(derived_data_path, '/tables')

# Mine clusters 
properties_clusters <- glue::glue(derived_tables_path, "/mine_properties_", cluster_type, ".geojson") %>% 
  sf::st_read(stringsAsFactors = FALSE) 
polygons_clusters <- glue::glue(derived_tables_path, "/mine_polygons_", cluster_type, ".geojson") %>% 
  sf::st_read(stringsAsFactors = FALSE) 

# Mine extraction 
production_clusters <- glue::glue(derived_tables_path, "/mine_production_", cluster_type, ".csv") %>% 
  readr::read_csv()
ore_processed_clusters <- glue::glue(derived_tables_path, "/mine_ore_processed_", cluster_type, ".csv") %>%
  readr::read_csv()

# Area intensity 
mine_area_intensity <- glue::glue(derived_tables_path, "/mine_area_intensity_",cluster_type,".csv") %>% 
  readr::read_csv()
```

## Mining properties X Development stage  
```{r properties-operation-stage}
properties_clusters %>%
  sf::st_drop_geometry() %>% 
  dplyr::group_by(development_group) %>% 
  dplyr::summarise(n = n()) %>% 
  dplyr::mutate(perc = n/sum(n)*100) 
```
*Note:* there is a large number of pre-operational properties, 73% of the properties. 


## Polygons cluster X Development stage  
```{r polygons-operation-stage}
polygons_clusters %>% 
  sf::st_drop_geometry() %>%
  tibble::as_tibble() %>% 
  dplyr::left_join(sf::st_drop_geometry(properties_clusters), by = c("hcluster_id" = "hcluster_id")) %>% 
  dplyr::mutate(development_group = ifelse(is.na(development_group), "Unknown", development_group)) %>% 
  dplyr::group_by(development_group) %>% 
  dplyr::summarise(n = n(), area = sum(area)) %>% 
  dplyr::mutate(n.perc = n/sum(n)*100, area.perc = area/sum(area)*100) 
```

*Notes:*
  - 240 (8.44%) from the polygons clusters identified in the satellite images are tagged pre-operational near to mining properties. This may indicate out-of-date information in the mining properties database. 
  - 309 (10.8%) of the polygons clusters were not linked to any mining property. In terms of area only 2,654 mk^2 (4.63%) were not linked to any mining property.
  - 115 (4.05%) from polygons clusters are closed. In terms of area only closed mines accounts only for 1.09%. This could indicate a bias in the mapping as closed area may have vegetation regeneration, which make identification more difficult. 
  - 89% of the polygons area is operational. Meaning that this areas may still expand in the near future and could be monitored from satellite. 

## Polygons cluster X commodities
```{r polygons-renewables}
polygons_clusters %>% 
  sf::st_drop_geometry() %>%
  tibble::as_tibble() %>% 
  dplyr::left_join(sf::st_drop_geometry(properties_clusters), by = c("hcluster_id" = "hcluster_id")) %>% 
  dplyr::mutate(commodities_list = ifelse(is.na(commodities_list), "Unknown", commodities_list)) %>% 
  dplyr::group_by(commodities_list) %>% 
  dplyr::summarise(n = n(), area = sum(area)) %>% 
  dplyr::mutate(n.perc = n/sum(n)*100, area.perc = area/sum(area)*100) %>% 
  dplyr::arrange(desc(area.perc)) %>% 
  dplyr::filter(cumsum(area.perc)<50)
```

## Commodity X Area (Muiti-counting)
```{r area-double-counting}
polygons_clusters %>% 
  sf::st_drop_geometry() %>%
  tibble::as_tibble() %>% 
  dplyr::left_join(sf::st_drop_geometry(properties_clusters), by = c("hcluster_id" = "hcluster_id")) %>% 
  dplyr::filter(!is.na(commodities_list)) %>% 
  dplyr::mutate(companionality = factor(stringr::str_detect(commodities_list, ","), c(T, F), c("Companion", "Single host"))) %>% 
  tidyr::separate_rows(commodities_list, sep = ",") %>% 
  dplyr::filter(commodities_list != "NA") %>% 
  group_by(commodities_list, companionality) %>% 
  dplyr::summarise(area = sum(area), .groups = 'drop') %>% 
  dplyr::group_by(commodities_list) %>% 
  dplyr::mutate(companionality_sum = sum(area)) %>%
  dplyr::ungroup() %>% 
  dplyr::arrange(dplyr::desc(companionality_sum), companionality) %>% 
  dplyr::filter(0.99 >= cumsum(area)/sum(area)) %>%
  dplyr::arrange(dplyr::desc(area)) %>% 
  ggplot(aes(x = reorder(commodities_list, companionality_sum), y = area, fill = companionality)) +
    geom_bar(stat = 'identity', position = 'stack') +
    coord_flip() +
    viridis::scale_fill_viridis(discrete = T, option = "E", begin = 0.8, end = 0.3) +
    scale_y_continuous(labels = scales::unit_format(unit = "k", scale = 1e-3, accuracy = 1)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = 'Commodities accounting for 99 % of the mining area',
       x = 'Commodity',
       y = 'Area [sqkm]') +
   guides(fill = guide_legend(title="Spatial cluster type"))
```
*Note:* Gold is extracted in most of the mapped areas 

## Commodity (production-allocated) X Area 
```{r area-allocated-commodity-production}
production_clusters %>% 
  dplyr::filter(2000 <= year, year <= 2019) %>% 
  dplyr::group_by(hcluster_id, commodity_name) %>% 
  dplyr::summarise(quantity = sum(quantity, na.rm = TRUE), quantity_unit = unique(quantity_unit), .groups = 'drop') %>% 
  dplyr::right_join(sf::st_drop_geometry(polygons_clusters), by = c("hcluster_id" = "hcluster_id")) %>% 
  dplyr::group_by(hcluster_id) %>% 
  dplyr::filter(commodity_name != "NA") %>% 
  dplyr::mutate(area = quantity/sum(quantity, na.rm = TRUE) * sum(area, na.rm = TRUE)) %>%
  dplyr::ungroup() %>% 
  dplyr::group_by(commodity_name) %>% 
  dplyr::summarise(area = sum(area, na.rm = TRUE), .groups = 'drop') %>% 
  dplyr::arrange(dplyr::desc(area)) %>% 
  dplyr::arrange(dplyr::desc(area)) %>% 
  ggplot(aes(x = reorder(commodity_name, area), y = area)) +
    geom_bar(stat = 'identity', position = 'stack') +
    coord_flip() +
    viridis::scale_fill_viridis(discrete = T, option = "E", begin = 0.8, end = 0.3) +
    scale_y_continuous(labels = scales::unit_format(unit = "k", scale = 1e-3, accuracy = 1)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = 'Mining area X Commodities',
       x = 'Commodity',
       y = 'Area [sqkm]') +
   guides(fill = guide_legend(title="Spatial cluster type"))
```
*Note* Allocated per production Coal is extracted in most of the mapped area, followed by Copper and Iron ore. Gold is only the fourth. 

## Commodity (production-allocated-split-companion) X Area 
```{r area-allocated-commodity-production-companion}
mine_area_intensity %>% 
  dplyr::group_by(commodity_name, companion) %>% 
  dplyr::summarise(area = sum(area, na.rm = TRUE), .groups = 'drop') %>% 
  dplyr::arrange(dplyr::desc(area)) %>% 
  ggplot(aes(x = reorder(commodity_name, area), y = area, fill = companion)) +
    geom_bar(stat = 'identity', position = 'stack') +
    coord_flip() +
    viridis::scale_fill_viridis(discrete = T, option = "E", begin = 0.8, end = 0.3) +
    scale_y_continuous(labels = scales::unit_format(unit = "k", scale = 1e-3, accuracy = 1)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = 'Mining area allocated to the commodities using production mass',
       x = 'Commodity',
       y = 'Area [sqkm]') +
   guides(fill = guide_legend(title="Spatial cluster type"))
```
*Note* This graph shows the shares of the area where a commodity is extracted as single host and the share of area where it is extracted with companions.  

## Land intensity table
```{r scater-production-area}
area_intensity_clusters %>% 
  ggplot(aes(x = quantity, y = area)) + 
  geom_point(size = 0.3) + 
  geom_smooth(method="lm") +
  facet_wrap(~commodity_name, scales = "free")
```

### Scarter plot production x area 
```{r scater-production-area}
area_intensity_clusters %>% 
  ggplot(aes(x = quantity, y = area)) + 
  geom_point(size = 0.3) + 
  geom_smooth(method="lm") +
  facet_wrap(~commodity_name, scales = "free")
```
*Note:* Several commodities show a nearly horizontal "flat" regression line, i.e. the same size of area can have very different reported production. This indicate that the same mineral could have different impacts in differences regions. 

## Mining land use per biome 
```{r area-mining-biome}
mine_polygons %>% 
  sf::st_drop_geometry() %>% 
  tibble::as_tibble() %>% 
  dplyr::group_by(biome_name) %>% 
  dplyr::summarise(area = sum(area, na.rm = TRUE), .groups = 'drop') %>% 
  dplyr::arrange(dplyr::desc(area)) %>% 
  ggplot(aes(x = reorder(biome_name, area), y = area)) +
    geom_bar(stat = 'identity', position = 'stack') +
    coord_flip() +
    viridis::scale_fill_viridis(discrete = T, option = "E", begin = 0.8, end = 0.3) +
    scale_y_continuous(labels = scales::unit_format(unit = "k", scale = 1e-3, accuracy = 1)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = 'Mining area per Biome',
       x = 'Commodity',
       y = 'Area [sqkm]') +
   guides(fill = guide_legend(title="Spatial cluster type"))
```

## TODO
```{r area-intensity-mining-biome}
mine_area_intensity %>% 
  dplyr::group_by(biome_name, commodity_name) %>% 
  dplyr::summarise(area = sum(area, na.rm = TRUE),
                   quantity = sum(area, na.rm = TRUE), .groups = 'drop') %>% 
  dplyr::mutate(area_intensity = area / quantity) %>% 
  dplyr::filter(area > 1000) %>% 
  dplyr::arrange(dplyr::desc(area_intensity)) %>% 
  ggplot(aes(x = reorder(biome_name, area_intensity), y = area_intensity)) +
    geom_bar(stat = 'identity', position = 'stack') +
    coord_flip() +
    facet_wrap(~commodity_name, scales = "free") + 
    viridis::scale_fill_viridis(discrete = T, option = "E", begin = 0.8, end = 0.3) +
    scale_y_continuous(labels = scales::unit_format(unit = "k", scale = 1e-3, accuracy = 1)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = 'Mining area intensity per Biome',
       x = 'Commodity',
       y = 'Area [sqkm]') +
   guides(fill = guide_legend(title="Spatial cluster type"))
```
